---
title: "Analysis species abundance data by iNEXT"
author: "T.C. Hsieh"
date: "Wednesday, August 27, 2014"
output: html_document
---

This is an simple guide for iNEXT. First install iNEXT R package from github repository (iNEXT is prepareing to submit to CRAN) by the following commends:

```{r}
# install.packages('devtools') # Tools to install R packages from github github repositories
devtools::install_github('iNEXT', 'JohnsonHsieh') # install iNEXT version 2.0 by devtools

```

Next, import iNEXT package and dataset

```{r}
library(iNEXT)
spec <- cbind(High =c(7,0,9,19,84,159),
              Intermediate =c(0,0,45,41,146,63),
              Low =c(15,3,3,10,126,4))
spec
```

Third, calculus species accumulation for each sites
```{r}
# Compute species accumulations
out <- iNEXT(x=spec, datatype="abundance", endpoint=200, knots=100, nboot=100)
# Data visualization
ggiNEXT(out)
ggiNEXT(out, type=1, facet.var = "site", color.var = "site")
ggiNEXT(out, type=1, color.var = "site")
ggiNEXT(out, type=2, color.var = "site")
ggiNEXT(out, type=3, color.var = "site")
```

Further, you can compute species diversity accumulations (e.g. Shannon entropy and Simpson index), see more introduction in [Chao et al. (2014)](http://www.esajournals.org/doi/abs/10.1890/13-0133.1)
```{r}
# Compute species accumulations
out <- iNEXT(x=spec, q=c(0,1,2), datatype="abundance", endpoint=200, knots=100, nboot=100)
# Data visualization
ggiNEXT(out, type=1, facet.var = "site", color.var = "order")
ggiNEXT(out, type=2, color.var = "site")
ggiNEXT(out, type=3, facet.var = "order", color.var = "both")
```

## How to cite this package
```{r}
# Use the following command
citation("iNEXT")
```

