---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

iNEXT (R package)
=====
<h4 style="text-align: right;">Most recent update time: January 8, 2014    

by Hsieh, T. C., K. H. Ma, and A. Chao    

Institute of Statistics, National Tsing Hua University, Hsin-Chu, Taiwan 30043</h4>


The program iNEXT (iNterpolation and EXTrapolation) is an updated R package from the original version supplied in Appendix of Chao et al. (2014). In this updated version, we have added more user-friendly features. Please notice the updated time and replace your old version by the most recent one.

The user provides a vector of abundances of individual species (abundance data) or incidences of individual species (incidence data). iNEXT computes the iNEXT computes the following species richness estimates and associated 95% confidence intervals:

1. Sample-size-based rarefaction and extrapolation: species richness estimates for rarefied and extrapolated samples up to a size specified by the user (i.e., an endpoint, see below). Refer to Colwell et al. (2012) for details.
2. Coverage-based rarefaction and extrapolation: species richness estimates for rarefied and extrapolated samples for sample coverage up to a coverage specified by the user (i.e., an endpoint, see below). Refer to Chao and Jost (2012) for details.

iNEXT also plots the following three integrated sampling curves suggested in Chao et al. (2014) for unified sampling and estimation in species diversity analysis.

- Sample-size-based rarefaction and extrapolation sampling curve: this curve plots the species richness estimates for a rarefied and extrapolated sample with respect to sample size up to a specified endpoint.
```{r eval=FALSE}
plot(object.iNEXT, type=1) # S3 method
ggiNEXT(object.iNEXT, type=1) # ggplot2 method
```

- Sample completeness curve: this curve plots the sample completeness (as measured by sample coverage) with respect to sample size. The curve provides a bridge between sample-size- and coverage-based rarefaction and extrapolation.
```{r eval=FALSE}
plot(object.iNEXT, type=2) # S3 method
ggiNEXT(object.iNEXT, type=2) # ggplot2 method
```

- Coverage-based rarefaction and extrapolation sampling curve: this curve plots the species richness estimates for rarefied sample and extrapolated sample with respect to sample coverage up to a specified endpoint.

```{r eval=FALSE}
plot(object.iNEXT, type=3) # S3 method
ggiNEXT(object.iNEXT, type=3) # ggplot2 method
```

### Software needed to run the development version

-   Required: [R](http://cran.rstudio.com/)
-   Suggested: [RStudio IDE](http://www.rstudio.com/ide/download/)

### How to run: start R(Studio) and copy-and-paste the following commands:

Start R(Studio) and copy-and-paste the commands below:
```{r install, eval=FALSE}
install.packages('devtools')
library(devtools)
install_github('JohnsonHsieh/iNEXT')
library(iNEXT)
```


Remark that in order to install `devtools` package, you should update R
to the last version. Further, to get `install_github` to work, you
should install the `httr` package.

### Example

Copy-and-paste the commands below:
```{r setup, include=FALSE}
knitr::opts_chunk$set(comment="", message=FALSE, warning=FALSE)
library(iNEXT)
library(ggplot2)
```

```{r eval=FALSE}
data(spider)
out <- iNEXT(spider, q=c(0, 1, 2), datatype="abundance", endpoint=500)
ggiNEXT(out, type=1, facet.var="site")
ggiNEXT(out, type=1, facet.var="order", color.var="site")
```
```{r ex1, echo=FALSE}
data(spider)
out <- iNEXT(spider, q=c(0, 1, 2), datatype="abundance", endpoint=500)
ggiNEXT(out, type=1, facet.var="site")
ggiNEXT(out, type=1, facet.var="order", color.var="site")
```


```{r eval=FALSE}
ggiNEXT(out, type=2, facet.var="none", color.var="site")
```
```{r ex2, echo=FALSE}
ggiNEXT(out, type=2, facet.var="none", color.var="site")
```

```{r eval=FALSE}
ggiNEXT(out, type=3, facet.var="site")
ggiNEXT(out, type=3, facet.var="order", color.var="site")
```
```{r ex3, echo=FALSE}
ggiNEXT(out, type=3, facet.var="site")
ggiNEXT(out, type=3, facet.var="order", color.var="site")

```

Run a sample-based example
```{r ex4, eval=FALSE}
data(ant)
iNEXT(ant$h50m, q=c(0,1,2), datatype="incidence") 
```



### How to cite

If you use iNEXT to obtain results for publication, you should cite at least one of the relevant papers (Chao and Jost 2012; Colwell et al. 2012; Chao et al. 2013) along with the following reference for iNEXT:

> Hsieh, T. C., K. H. Ma, and A. Chao. 2014. iNEXT: An R package for interpolation and extrapolation in measuring species diversity. Unpublished manuscript.

### License

The iNEXT package is licensed under the GPLv2. To help refine iNEXT, your comments or feedbacks would be welcome (please send them to [Anne Chao](chao@stat.nthu.edu.tw)).

### Referance

1. Chao, A., N. J. Gotelli, T. C. Hsieh, E. L. Sander, K. H. Ma, R. K. Colwell, and A. M. Ellison 2014. Rarefaction and extrapolation with Hill numbers: a unified framework for sampling and estimation in biodiversity studies, Ecological Monographs 84:45-67.
2. Chao, A., and L. Jost. 2012. Coverage-based rarefaction and extrapolation: standardizing samples by completeness rather than size. Ecology 93:2533-2547.
3. Colwell, R. K., A. Chao, N. J. Gotelli, S. Y. Lin, C. X. Mao, R. L. Chazdon, and J. T. Longino. 2012. Models and estimators linking individual-based and sample-based rarefaction, extrapolation and comparison of assemblages. Journal of Plant Ecology 5:3-21.
4. Ellison, A. M., A. A. Barker-Plotkin, D. R. Foster, and D. A. Orwig. 2010. Experimentally testing the role of foundation species in forests: the Harvard Forest Hemlock Removal Experiment. Methods in Ecology and Evolution 1:168-179.
5. Hsieh, T. C., K. H. Ma, and A. Chao. 2013. iNEXT online: interpolation and extrapolation (Version 1.3.0) [Software]. Available from http://chao.stat.nthu.edu.tw/blog/software-download/.
6. Hsieh T. C., K. H. Ma, and A. Chao. 2014. iNEXT: An R package for interpolation and extrapolation in measuring species diversity. Unpublished manuscript.
7. Longino, J. T., and R. K. Colwell. 2011. Density compensation, species composition, and richness of ants on a neotropical elevational gradient. Ecosphere 2:art29.
8. Sackett, T. E., S. Record, S. Bewick, B. Baiser, N. J. Sanders, and A. M. Ellison. 2011. Response of macroarthropod assemblages to the loss of hemlock (Tsuga canadensis), a foundation species. Ecosphere 2: art74.
