---
title: "Customize Plots and Graphics"
author: "T. C. Hsieh, K. H. Ma, and Anne Chao"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
vignette: >
      %\VignetteIndexEntry{Customize Plots and Graphics}
      %\VignetteEngine{knitr::rmarkdown}
      %\VignetteEncoding{UTF-8}
---
```{r include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "", 
                      fig.retina=2,
                      fig.align='center',
                      fig.width = 8, fig.height = 6,
                      out.width = "100%")
library(BiocStyle)
```


## General customization
The data visualization package `r CRANpkg("ggplot2")` provides `scale_` function to customize  data which is mapped into an aesthetic property of a `geom_`. The following functions would help user to customize `ggiNEXT` output.

- change point shape: `scale_shape_manual`    
- change line type : `scale_linetype_manual`    
- change line color: `scale_colour_manual`    
- change band color: `scale_fill_manual`    
see [quick reference](http://sape.inf.usi.ch/quick-reference/ggplot2/scale) for style setting.

### Example: `spider` data
To show how to custmized `ggiNEXT` output, we use abundance-based data `spider` as an example.
```{r}
library(iNEXT)
library(ggplot2)
library(gridExtra)
library(grid)
data("spider")
out <- iNEXT(spider, q=0, datatype="abundance")
g <- ggiNEXT(out, type=1, color.var = "site")
g
```

### Change shapes, line types and colors
```{r}
g1 <- g + scale_shape_manual(values=c(11, 12)) + 
          scale_linetype_manual(values=c(1,2))
g2 <- g + scale_colour_manual(values=c("red", "blue")) +
          scale_fill_manual(values=c("red", "blue"))

# Draw multiple graphical objec on a page
# library(gridExtra)
grid.arrange(g1, g2, ncol=2)
```

## Customize point/line size by hacking
In order to chage the size of reference sample point or rarefaction/extrapolation curve, user need modify `ggplot` object.

- change point size:    
the reference sample size point is drawn on the first layer by `ggiNEXT`. Hacking point size by the following
```{r}
# point is drawn on the 1st layer, default size is 5
gb3 <- ggplot_build(g)
gb3$data[[1]]$size <- 10
gt3 <- ggplot_gtable(gb3)

# use grid.draw to draw the graphical object
# library(grid)
# grid.draw(gt3)
```

- change line width (size):      
the reference sample size point is drawn on the second layer by `ggiNEXT`. Hacking point size by the following
```{r}
# line is drawn on the 2nd layer, default size is 1.5
gb4 <- ggplot_build(g)
gb4$data[[2]]$size <- 3
gt4 <- ggplot_gtable(gb4)
# grid.draw(gt4)
```
```{r}
grid.arrange(gt3, gt4, ncol=2)
```


## Customize theme
A `ggplot` object can be themed by adding a theme. User could run `help(theme_grey)` to show the default themes in `ggplot2`. Further, some extra themes provided by `r CRANpkg("ggthemes")` package. Examples shown in the following:
```{r}
g5 <- g + theme_bw() + theme(legend.position = "bottom")
g6 <- g + theme_classic() + theme(legend.position = "bottom")
grid.arrange(g5, g6, ncol=2)
```
```{r}
library(ggthemes)
g7 <- g + theme_hc(bgcolor = "darkunica") +
          scale_colour_hc("darkunica")

g8 <- g + theme_economist() + scale_colour_economist()
grid.arrange(g7, g8, ncol=2)
```

### Black-White theme
The following are custmized themes for black-white figure. To modifiy legend, see [Cookbook for R](http://www.cookbook-r.com/Graphs/Legends_(ggplot2)/) for more details.

```{r}
g9 <- g + theme_bw(base_size = 18) +
      scale_fill_grey(start = 0, end = .4) +
      scale_colour_grey(start = .2, end = .2) +
      theme(legend.position="bottom",
            legend.title=element_blank())

g10 <- g + theme_tufte(base_size = 12) +       
    scale_fill_grey(start = 0, end = .4) +
    scale_colour_grey(start = .2, end = .2) +
    theme(legend.position="bottom",
          legend.title=element_blank())
grid.arrange(g9, g10, ncol=2)
```

### Draw R/E curves by yourself
In `r CRANpkg("iNEXT")`, we provide a S3 `ggplot2::fortify` method for class `iNEXT`. The function `fortify` offers a single plotting interface for rarefaction/extrapolation curves. Set  argument `type = 1, 2, 3` to plot the corresponding rarefaction/extrapolation curves.
```{r}
df <- fortify(out, type=1)
head(df)

df.point <- df[which(df$method=="observed"),]
df.line <- df[which(df$method!="observed"),]
df.line$method <- factor(df.line$method, 
                         c("interpolated", "extrapolated"),
                         c("interpolation", "extrapolation"))
 
ggplot(df, aes(x=x, y=y, colour=site)) + 
  geom_point(aes(shape=site), size=5, data=df.point) +
  geom_line(aes(linetype=method), lwd=1.5, data=df.line) +
  geom_ribbon(aes(ymin=y.lwr, ymax=y.upr,
                  fill=site, colour=NULL), alpha=0.2) +
  labs(x="Number of sampling units", y="Species diversity") +
  theme(legend.position = "bottom", 
        legend.title=element_blank(),
        text=element_text(size=18)) 
```

